---
title: "STA130H1F -- Winter 2021"
author: NAME and Student ID
subtitle: Week 5 Problem Set
output:
  pdf_document: default
---


```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE)
```

# Instructions

## How do I hand in these problems for the Thursday deadline?

Your complete .Rmd file that you create for these practice problems AND the resulting .pdf (i.e., the one you 'Knit to PDF' from your .Rmd file) must be uploaded into a Quercus assignment (link: [https://q.utoronto.ca/courses/253019/assignments/801314](https://q.utoronto.ca/courses/253019/assignments/801314) by 11:59 a.m. ET (midday) on February 10, 2022. Late problem sets or problems submitted another way (e.g., by email) are *not* accepted.


# Problem set grading

There are two parts to your problem set. Part one is R based: there are 2 mandatory questions in part one, and additional optional questions. Optional R based questions are highly recommended as they will help prepare you for larger assessments (and help improve your R skills, of course!); but, they are *not* part of the grading for the weekly problem sets. Part two is (more) writing based: all of part two is mandatory. We recommend you use a word processing software like Microsoft Word to check for grammar errors in your written work. Note: there can be issues copying from Word to R Markdown so it may be easier to write in this file first and then copy the text to Word. Then you can make any changes flagged in Word directly in this file.
  
\newpage 

# Part 1

## Question 1 [Required] 

### There have been many questions regarding whether or not usage of social media increases anxiety levels. A study was conducted to examine the relationship between social media usage and student anxiety. The following scores are measures of anxiety levels of students on a Monday. Higher scores indicate higher anxiety; and, if a student used social media more than 1 hour per day then their usage was categorized as "High".

```{r}
social_media_usage <- c(rep("Low", 30), rep("High", 16));
anxiety_score <- c(24.64, 39.29, 16.32, 32.83, 28.02, 
                   33.31, 20.60, 21.13, 26.69, 28.90,
                   26.43, 24.23, 7.10,  32.86, 21.06,
                   28.89, 28.71, 31.73, 30.02, 21.96,
                   25.49, 38.81, 27.85, 30.29, 30.72,
                   21.43, 22.24, 11.12, 30.86, 19.92,
                   33.57, 34.09, 27.63, 31.26,
                   35.91, 26.68, 29.49, 35.32,
                   26.24, 32.34, 31.34, 33.53,
                   27.62, 42.91, 30.20, 32.54)

anxiety_data <- tibble(social_media_usage, anxiety_score)
glimpse(anxiety_data)
```


### (a) Construct boxplots of `anxiety_score` for the two levels of social media usage. Write 2-3 sentences comparing the distributions of anxiety scores across the social media usage groups.

```{r, fig.height=2, fig.width=3.5}
```


### (b) Do these data support the claim that the median anxiety level is different for those who use social media in high frequency compared to those who use social media in lower frequency? 

(i) State the hypotheses you are testing (be sure to define any parameters you refer to).



(ii) Look at the code below and write a few sentences explaining what the code inside the for loop is doing and why.

```{r}
# Note: including the .groups="drop" option in summarise() will suppress a friendly 
# warning R prints otherwise "`summarise()` ungrouping output (override with 
#`.groups` argument)".  
  # Including the .groups="drop" option is optional, but you should include it if you 
#don't want to see that warning.
test_stat <- anxiety_data %>% group_by(social_media_usage) %>%
  summarise(medians = median(anxiety_score), .groups="drop") %>%
  summarise(value = diff(medians))
test_stat <- as.numeric(test_stat)
test_stat

set.seed(523)
repetitions <- 1000;
simulated_values <- rep(NA, repetitions)

for(i in 1:repetitions){
  simdata <- anxiety_data %>% mutate(social_media_usage = sample(social_media_usage))

  sim_value <- simdata %>% group_by(social_media_usage) %>%
    summarise(medians = median(anxiety_score), .groups="drop") %>%
    summarise(value = diff(medians))
  
  simulated_values[i] <- as.numeric(sim_value)
}

sim <- tibble(median_diff = simulated_values)

sim %>% ggplot(aes(x=median_diff)) + geom_histogram(binwidth=1, color="black", fill="gray")

# Calculate p-value
num_more_extreme <- sim %>% filter(abs(median_diff) >= abs(test_stat)) %>% summarise(n())

p_value <- as.numeric(num_more_extreme / repetitions)
p_value
```


(iii) Write a few sentences summarizing your conclusions. Be sure to interpret the p-value carefully and to clearly address the research question.


\newpage
## Question 2 [Optional] 

A criminal court considers two opposing claims about a defendant: they are either innocent or guilty. In the Canadian legal system, the role of the prosecutor is to present convincing evidence that the defendant is not innocent.  Lawyers for the defendant attempt to argue that the evidence is *not convincing* enough to rule out that the defendant could be innocent.  If there is not enough evidence to convict the defendant and they are set free, the judge generally does not deliver a verdict of "innocent", but rather of "not guilty".  

### (a) If we look at the criminal trial example in the hypothesis test framework, which would be the null hypothesis and which the alternative?


### (b) In the context of this problem, describe what rejecting the null hypothesis would mean.



### (c) In the context of this problem, describe what failing to reject the null hypothesis would mean.


### (d) In the context of this problem, describe what a type II error would be.


### (e)  In the context of this problem, describe what a type I error would be.




\newpage
## Question 3 [Optional]  
**(Adapted from "Biostatistics for the Biological and Health Sciences") **
The table below presents data from a random sample of passengers sitting in the front seat of cars involved in car crashes. Researchers are interested in whether the fatality rates (i.e. death rates) differ for passengers in cars with airbags and passengers in cars without airbags.

|                           | Airbag available | No airbag available |
|---------------------------|------------------|---------------------|                           
| Passenger Fatalities      |  45              | 62                  |
| Total number of Passengers|  10,541          | 9,867               |

The code below creates a tidy data frame for this problem, using the R command `rep`. This function creates a vector which replicates its first argument the number of times indicated by its second argument. For example, the `rep("hello", 5)` creates a vector with 5 elements, each of which is "hello".  Run the code chunk below to load a tidy tibble called `data` which you'll use for the remainder of this question.

```{r}
library(tidyverse)

data <- tibble(group=c(rep("airbag",10541),rep("no_airbag",9867)),
               outcome=c(rep("dead",45), rep("alive",10541-45), 
                         rep("dead",62), rep("alive",9867-62)))
```

### (a) State appropriate hypotheses to compare the proportions of deaths in cars with and without airbags. Be sure to define any parameters you refer to in your hypotheses.




### (b) Carry out a hypothesis test for the hypotheses stated in part (a).

```{r}
set.seed(523) # Replace the number in the parentheses with the 1st, 3rd, and 5th
# digits in your student number.

```

```{r}
# space for scratch work if needed

```


### (c) Based on your answer in part (b), would you reject the null hypothesis at the 0.1 significance level?


### (d) Based on your answer in part (c), what kind of error did you possibly make?


\newpage
### Question 4 [Optional] 

In class we've talked about two kinds of hypothesis tests. In the first kind (week 4) we talked about how to test whether a proportion is equal to a specific value, with hypotheses of the form: $H_0: p = p_0$ vs $H_A: p \neq p_0$.  In this week's class (week 5), we talked about how to test if there is a difference between two groups (e.g. a difference in the means of two groups, the medians of two groups, or proportions of two groups). A test for the difference between the means of two groups takes the form: $H_0: \mu_1 = \mu_2$ vs $H_A: \mu_1 \neq \mu_2$.

For each of the following scenarios, state appropriate hypotheses $H_0$ and $H_A$.  Be sure to carefully define any parameters you refer to.

### (a) A health survey asked individuals to report the number of times they exercised each week. Researchers were interested in determining if the proportion of individuals who exercised at least 100 minutes per week differed between people who live in the condos vs people who do not live in condos.


### (b) A study was conducted to examine whether a baby is born prematurely/early (i.e., before their due date) to whether or not the baby's mother smoked while she was pregnant.



### (c) Nintendo is interested in whether or not their online advertisements are working. They record whether or not a user had seen an ad on a given day and their amount of spending on Nintendo products in the next 48 hours. They are interested in determining if there is an association between whether or not the user say an ad and their expenditures.



### (d) Based on results from a survey of graduates from the University of Toronto, we would like to compare the median salaries of graduates from the statistics and graduates of mathematics programs.


\newpage

# Part 2 [Required]

# Part 2 (Choice of oral or written submission)

For this week, you can complete the required task as a written
assignment, OR you can submit an oral response (i.e. just sound, or
sound and video). While it is optional *this time*, there will be two
other problem set part \#2s that you will be required to complete an
oral response, and therefore it is a good idea to try and practice it
now!

Suppose you were given the following dataset and told it was a survey
conducted with a random sample of 12 student's in Prof Caetano's 4th year
class on statistical consultation, communication, and collaboration
(STA490). They were asked if they were currently living in Toronto
(defined as Toronto, Etobicoke, Scarborough, York, North York and East
York) or not and how long it took them to travel to the place they/their
family usually purchases groceries, using whatever travel method they
usually used to get there. **Student** is just an ID they were randomly
assigned when the data was anonymized.

| Student | Location    | Travel time (mins) |
|---------|-------------|--------------------|
| 113     | Toronto     | 20                 |
| 120     | Not Toronto | 17                 |
| 499     | Not Toronto | 35                 |
| 703     | Toronto     | 7                  |
| 190     | Toronto     | 12                 |
| 286     | Toronto     | 45                 |
| 203     | Not Toronto | 45                 |
| 503     | Toronto     | 15                 |
| 154     | Not Toronto | 50                 |
| 334     | Toronto     | 25                 |
| 417     | Not Toronto | 22                 |
| 303     | Toronto     | 20                 |

In this writing activity you will practice writing instructions and
demonstrate your understanding of simulation based hypothesis tests, the
focus of Module 4.

### Your task

Write instructions so that someone could perform ONE of the following
hypothesis tests, using only the resources listed below. You do not have
to use all the materials.

A)  Are 75% of the students taking STA490 this semester living in
    Toronto?
B)  Do students who live in Toronto have the same average travel time to
    their local supermarket as student who don't live in Toronto?

#### Resources

-   A pack of 52 playing cards, half red, half blue
-   A reusable shopping back
-   A friend who loves doing mental arithmetic and is very good at it
-   A printed sheet with all the travel times listed in a column, but
    nothing else
-   A white board
-   A whiteboard marker
-   A white board eraser

#### Key things to include

-   Briefly introduce which test you are choosing from above (A or B).
-   Describe the hypotheses for your chosen test in plain words.
-   Briefly describe why we would want to create a simulated sampling
    distribution.
-   Provide instructions that would enable a physical simulation of a
    sampling distribution for your chosen test---**using only the
    resources listed above**. Your final instruction should be "And
    repeat steps X to Y 99 more times to get 100 total repetitions".
    Replace X and Y with whatever the relevant steps are.
-   *Include **at least 2** vocabulary words from this module and
    explain/connect them with what you're doing in this physical
    simulation.*

## Some things to keep in mind

-   Try to not spend more than 20 minutes on the prompt.
-   Aim for more than 200 but less than 350 words (if **written**) and
    no more than 4 minutes (if **oral**)
-   Use full sentences (writing or speaking).
-   Grammar is not the main focus of the assessment, but it is important
    that you communicate in a clear and professional manner (i.e., no
    slang).
-   You do NOT need to have the resources (cards, etc.) above to do this
    task. There is no requirement to use them or actually carry out your
    instructions, just to create an appropriate instruction set.
-   Be specific. A good principle when responding to a writing prompt in
    STA130 is to assume that your audience is not aware of the subject
    matter (or in this case has not read the prompt). You would want
    another first year university students to be able to follow your
    instructions.

## Vocabulary

-   Statistical inference
-   Population
-   Random sample
-   Sampling distribution
-   Simulation
-   Parameter
-   Simulation statistic
-   Test statistic
-   P-value
-   Type 1 and 2 errors (also written as Type I and II errors)
-   Comparing two population means/medians/proportions
-   One- and two-group hypothesis tests
-   for() loops
-   sample()
-   diff()

## For those who choose to do the (optional) oral submission

- Provide an oral response (either voice clip or video) that is no more than 4 minutes in length. Do not feel the need to do tons of ‘takes’. Rather, you can repeat yourself if you make a mistake, or feel you are unclear. This is not meant to be an additional burden, but rather to provide you with the opportunity to practice your oral communication skills and get a break from writing.

- You might be wondering how can I record this? One way to do this would be to schedule a Zoom meeting and record yourself in it. You can "Record the video to the cloud" (or directly on your computer). Several file types will be saved, including a video version, and one that is just a voice recording.  If you are looking for more ideas of how to record yourself for this assignment or run into issues on how to upload your assignment, please post to Piazza.
 
- You MUST upload a link (aka a URL) to your recording for your TA to watch your video. **Paste the URL directly into your answers to this problem set.** ONLY links will be reviewed by the TA, we are not accepting mp4 or clip uploads. You can provide the zoom cloud link from your recording if you've chosen "Record to the cloud". Or, alternatively, you can upload your video to [MyMedia](https://mymedia.library.utoronto.ca), MS Stream, YouTube, etc., though uploading/processing can be slow, so give yourself plenty of time to prepare this.


  - NOTE: Your link must NOT be password protected (which is often the default in Zoom sharing). Please ensure that there is NO password protection on the video/link. You can test this by seeing if you can open the file in a new browser or in incognito mode. [**Files which cannot be accessed by the TA just "clicking the link" and "watching the video" (without entering a password) will receive a 0.**]{.ul}
  
  - There is more information about Zoom sharing settings [here](https://support.zoom.us/hc/en-us/articles/205347605-Managing-and-sharing-cloud-recordings).

## Example

For example, include a line, such as the following, but change the link to your own video link:

**URL for my submission:** <https://www.youtube.com/watch?v=dQw4w9WgXcQ>.

This is an acceptable submission because:

- There is a link directly in your Rmd and pdf submission.
- The link is not a password/passcode needed to watch/listen to the video.


